## This file contains the code used to construct the corpus. Further down in the script is code written to perform basic exploration of the data, none of which is included in the final paper. 

```{r}
library("dplyr")
library(quanteda)
library(readtext)
library(stringr)
library("quanteda.textstats")
library(purrr)
library(rsvd)
library(Rtsne)
library(geometry)
library(stm)
```

# PROCESS THE RAW TXT FILES
```{r}
# Build an ARTICLE-LEVEL corpus from Factiva .txt exports

data_path <- "/Users/imanfq/Desktop/ASDS/Dissertation/data"

files <- list.files(data_path, pattern = "\\.txt$",
                    recursive = TRUE, full.names = TRUE)
rt <- readtext(files, encoding = "UTF-8")   

rel    <- sub(paste0("^", data_path, "/"), "", files)   
parts  <- strsplit(rel, "/")
f_country <- vapply(parts, `[`, character(1), 1)         
f_period  <- vapply(parts, `[`, character(1), 2)         
folder    <- vapply(parts, `[`, character(1), 3)         
f_outlet  <- sub("(Pre|Post)$", "", folder)              

FACTIVA_ID <- "Document [A-Z0-9]{2,12}20[0-9]{6}[a-z0-9]{3,}"
DOC_LINE   <- paste0("(\\r?\\n", FACTIVA_ID, ")(?=\\r?\\n|$)")

split_articles <- function(txt) {
  marked <- str_replace_all(txt, DOC_LINE, "\\1<<<SPLIT>>>")
  pcs <- str_split(marked, fixed("<<<SPLIT>>>"))[[1]]
  pcs <- str_trim(pcs)
  pcs[str_detect(pcs, FACTIVA_ID)]
}

art_list <- Map(function(txt, ctry, per, out, src) {
  pcs <- split_articles(txt)
  if (length(pcs) == 0) return(NULL)
  data.frame(
    text       = pcs,
    country    = ctry,
    period     = per,
    outlet     = out,
    srcfile    = src,
    article_n  = seq_along(pcs),
    factiva_id = str_match(pcs, "Document ([A-Z0-9]{2,12}20[0-9]{6}[a-z0-9]{3,})")[, 2],
    stringsAsFactors = FALSE
  )
}, rt$text, f_country, f_period, f_outlet, rel)

names(art_list) <- NULL
art_df <- do.call(rbind, art_list) 
rownames(art_df) <- NULL


dups <- duplicated(art_df$factiva_id)
cat("Articles parsed:", nrow(art_df),
    "| duplicates removed:", sum(dups), "\n")
art_df <- art_df[!dups, ]


art_df$doc_id <- paste(art_df$outlet, art_df$period, art_df$factiva_id, sep = "_")
```


# CLEANING

```{r}
#Remove copyright block and maintain headline 
library(stringr)

parse_factiva <- function(txt) {
  
  bits   <- str_split(txt, regex("(?:opyright|©)[^\n]*\n", ignore_case = TRUE), n = 2)[[1]]
  header <- bits[1]
  body   <- if (length(bits) > 1) bits[2] else ""


  pre_wc      <- str_split(header, regex("\n\\s*\\d+\\s+words\\b"), n = 2)[[1]][1]
  first_block <- str_split(str_trim(pre_wc), regex("\n\\s*\n"))[[1]][1]
  hl_lines    <- str_split(str_trim(first_block), "\n")[[1]]

  headline    <- if (length(hl_lines) >= 2) str_c(hl_lines[-1], collapse = " ") else hl_lines[1]

  body <- body |>
    str_remove(regex("(?s)\\s*Document\\s+\\S+\\s*$")) |>   
    str_squish()

  list(headline = str_squish(headline), body = body)
}

article_text = art_df$text 

parsed <- purrr::map(article_text, parse_factiva)

art_df$headline   <- purrr::map_chr(parsed, "headline")
art_df$body       <- purrr::map_chr(parsed, "body")



art_df$year <- as.integer(str_sub(
  str_match(art_df$factiva_id, "20[0-9]{6}")[, 1], 1, 4
))


art_df$year_text <- as.integer(str_extract(art_df$text, "\\b(19|20)\\d{2}\\b"))
art_df$year_mismatch <- art_df$year != art_df$year_text & !is.na(art_df$year_text)

cat("Rows missing a year in raw text:", sum(is.na(art_df$year_text)), "\n")
cat("Rows where accession year vs text year disagree:", sum(art_df$year_mismatch, na.rm = TRUE), "\n")


art_df$years_since_chatgpt <- art_df$year - 2022

protect_countries <- function(x) {
  x <- str_replace_all(x, "\\bUnited States\\b", "USCOUNTRY")
  x <- str_replace_all(x, "\\bU\\.S\\.A\\.?",     "USCOUNTRY")
  x <- str_replace_all(x, "\\bU\\.S\\.",          "USCOUNTRY")
  x <- str_replace_all(x, "\\bUSA\\b",            "USCOUNTRY")
  x <- str_replace_all(x, "\\bUS\\b",             "USCOUNTRY")
  x <- str_replace_all(x, "\\bUnited Kingdom\\b", "UKCOUNTRY")
  x <- str_replace_all(x, "\\bU\\.K\\.",          "UKCOUNTRY")
  x <- str_replace_all(x, "\\bUK\\b",             "UKCOUNTRY")
  x
}

art_df$text_clean <- protect_countries(str_c(art_df$headline, art_df$body, sep = ". "))
```

```{r}
art_df <- art_df |>
  select(-year_text, -year_mismatch)
```

```{r}
bar_dat = art_df |> group_by(period, outlet) |> summarise(count = n())

bar_dat %>%
   mutate(
     outlet = factor(outlet,
                     levels = c("FT", "Guardian", "Sun",
                               "WSJ", "NYTimes", "NP")),
     period = factor(period, levels = c("Pre", "Post"))
   ) %>%
   ggplot(aes(x = outlet, y = count, fill = period)) +
   geom_col(position = position_dodge(width = 0.8), width = 0.7) +
   geom_text(aes(label = count),
             position = position_dodge(width = 0.8),
             vjust = -0.4, size = 3) +
   scale_fill_manual(values = c("Pre" = "grey65", "Post" = "steelblue4")) +
   labs(
     title = "Article Volume by Outlet, Before and After ChatGPT's Release",
     subtitle = "UK outlets (FT, Guardian, Sun) and US outlets (WSJ, NYTimes, NY Post)",
     x = NULL, y = "Number of articles", fill = "Period"
   ) +
   theme_minimal(base_size = 11) +
   theme(
     panel.grid.major.x = element_blank(),
     legend.position = "top"
   )
```


# CONVERT TO CORPUS
```{r}
# CORPUS
full_corp <- art_df |> select(doc_id, text_clean, headline, body, country, period, outlet, factiva_id, srcfile) |>
  corpus(text_field = "text_clean", docid_field = "doc_id")
```

# TOKENIZE AND REMOVE STOPWORDS AND STEM

```{r}
#solving for collocations and then tokenizing and removing stop words and stemming
sw <- stopwords("english")

toks_coll <- full_corp |>
  tokens(remove_punct = TRUE, remove_symbols = TRUE, remove_numbers = TRUE) |>
  tokens_tolower()

edge_stop <- function(x) {
  w <- strsplit(x, " ")[[1]]
  w[1] %in% sw || w[length(w)] %in% sw
}
```

```{r}
colls <- textstat_collocations(toks_coll, size = 2:3, min_count = 25)
keep  <- colls[colls$lambda > 5 & !vapply(colls$collocation, edge_stop, logical(1)), ]
keep <- keep[order(-lengths(strsplit(keep$collocation, " "))), ]   # longest first

nrow(keep)   
```

```{r}
extra_stops <- c("said", "say", "says", "told", "according",
                 "use", "can", "one") 

full_tokenization <- toks_coll |>
  tokens_compound(pattern = phrase(keep$collocation), concatenator = "_") |> tokens_remove(extra_stops) |>
  tokens_remove(stopwords("english"))
```


```{r}
full_tokenization = readRDS("full_tokenization.rds")
```

# SUBSET BY COUNTRY AND PERIOD
```{r}
uk_pre_toks  <- tokens_subset(full_tokenization, country == "UK" & period == "Pre")
uk_post_toks  <- tokens_subset(full_tokenization, country == "UK" & period == "Post")

us_pre_toks  <- tokens_subset(full_tokenization, country == "US" & period == "Pre")
us_post_toks  <- tokens_subset(full_tokenization, country == "US" & period == "Post")
```

# DFMs

```{r}
#note: secondlast seminar of my472 has this code 
dfm_full = full_tokenization |> dfm(tolower = TRUE,
      remove_padding = FALSE, 
    verbose = quanteda_options("verbose"))

dfm_ukpre = uk_pre_toks |> dfm(tolower = TRUE, 
      remove_padding = FALSE, 
    verbose = quanteda_options("verbose"))


dfm_ukpost = uk_post_toks |> dfm(tolower = TRUE,
      remove_padding = FALSE, 
    verbose = quanteda_options("verbose"))


dfm_uspre = us_pre_toks |> dfm(tolower = TRUE,
      remove_padding = FALSE, 
    verbose = quanteda_options("verbose"))

dfm_uspost = us_post_toks |> dfm(tolower = TRUE, 
      remove_padding = FALSE, 
    verbose = quanteda_options("verbose"))

```

## Exploratory 

```{r}
outlet_counts = article_df |> group_by(outlet) |> summarise(count = n())
print(outlet_counts)

country_period_counts = article_df |> count(country, period, outlet) 
print(country_period_counts)

country_period_counts_vis = article_df %>%
  count(country, period, outlet) %>%
  mutate(period = factor(period, levels = c("Pre", "Post"))) %>%
  ggplot(aes(x = outlet, y = n, fill = period)) +
  geom_col(position = "dodge") +
  scale_y_log10() +
  facet_wrap(~country, scales = "free_x") +
  labs(y = "Number of Articles", x = NULL) +
  theme_minimal()

library(dplyr)
library(tidyr)

pct_change <- country_period_counts |>
  pivot_wider(names_from = period, values_from = n) |>
  mutate(
    log_ratio  = log(Post / Pre)
  )

country_period_pct_vis = pct_change %>%
  ggplot(aes(x = outlet, y = log_ratio)) +
  geom_col(position = "dodge") +
  scale_y_log10() +
  facet_wrap(~country, scales = "free_x") +
  labs(y = "Number of Articles", x = NULL) +
  theme_minimal()

print(country_period_pct_vis)
```


```{r}
#| eval: false

#convert to corpus object 

full_corp <- article_df |> select(doc_id, text_clean, headline, body, country, period, outlet, factiva_id, srcfile) |>
  corpus(text_field = "text_clean", docid_field = "doc_id")


sw <- stopwords("english")

toks_coll <- full_corp |>
  tokens(remove_punct = TRUE, remove_symbols = TRUE, remove_numbers = TRUE) |>
  tokens_tolower()

edge_stop <- function(x) {
  w <- strsplit(x, " ")[[1]]
  w[1] %in% sw || w[length(w)] %in% sw
}

colls <- textstat_collocations(toks_coll, size = 2:3, min_count = 25)
keep  <- colls[colls$lambda > 5 & !vapply(colls$collocation, edge_stop, logical(1)), ]
keep <- keep[order(-lengths(strsplit(keep$collocation, " "))), ]   # longest first  

extra_stops <- c("said", "say", "says", "told", "according",
                 "use", "can", "one") 

full_tokenization <- toks_coll |>
  tokens_compound(pattern = phrase(keep$collocation), concatenator = "_") |> tokens_remove(extra_stops) |>
  tokens_remove(stopwords("english")) |> tokens_remove(pattern = c("a.i", "AI", "artificial_intelligence"), valuetype = "fixed", case_insensitive = TRUE)

docvars(full_tokenization, "lang") <- cld2::detect_language(sapply(as.list(full_tokenization), paste, collapse = " "))
full_tokenization_en <- tokens_subset(full_tokenization, lang == "en" | is.na(lang)) 
```

```{r}
full_tokenization_en = readRDS("")
```


```{r}

ft_pre_toks        <- tokens_subset(full_tokenization_en, country == "UK" & outlet == "FT"       & period == "Pre")
ft_post_toks       <- tokens_subset(full_tokenization_en, country == "UK" & outlet == "FT"       & period == "Post")

guardian_pre_toks  <- tokens_subset(full_tokenization_en, country == "UK" & outlet == "Guardian" & period == "Pre")
guardian_post_toks <- tokens_subset(full_tokenization_en, country == "UK" & outlet == "Guardian" & period == "Post")

sun_pre_toks       <- tokens_subset(full_tokenization_en, country == "UK" & outlet == "Sun"      & period == "Pre")
sun_post_toks      <- tokens_subset(full_tokenization_en, country == "UK" & outlet == "Sun"      & period == "Post")


np_pre_toks        <- tokens_subset(full_tokenization_en, country == "US" & outlet == "NP"       & period == "Pre")
np_post_toks       <- tokens_subset(full_tokenization_en, country == "US" & outlet == "NP"       & period == "Post")

nytimes_pre_toks   <- tokens_subset(full_tokenization_en, country == "US" & outlet == "NYTimes"  & period == "Pre")
nytimes_post_toks  <- tokens_subset(full_tokenization_en, country == "US" & outlet == "NYTimes"  & period == "Post")

wsj_pre_toks       <- tokens_subset(full_tokenization_en, country == "US" & outlet == "WSJ"      & period == "Pre")
wsj_post_toks      <- tokens_subset(full_tokenization_en, country == "US" & outlet == "WSJ"      & period == "Post")
```

## Word Clouds 

## WSJ vs FT
```{r}
library(quanteda.textplots)
wsj_pre_toks |> dfm(tolower = TRUE, # yes, we want to lower case every token!
      remove_padding = FALSE, # keep default, but doesn't matter since we didn't add padding when tokenizing
    verbose = quanteda_options("verbose")) |> # keep default
      #color is for least to most frequent
   textplot_wordcloud(max_words = 50, color = c("red", "darkgreen", "blue")); title("WSJ: Pre")
```

```{r}
wsj_post_toks |> dfm(tolower = TRUE, # yes, we want to lower case every token!
      remove_padding = FALSE, # keep default, but doesn't matter since we didn't add padding when tokenizing
    verbose = quanteda_options("verbose")) |> # keep default
      #color is for least to most frequent
   textplot_wordcloud(max_words = 50, color = c("red", "darkgreen", "blue")); title("WSJ: Post")
```

```{r}
ft_pre_toks |> dfm(tolower = TRUE, # yes, we want to lower case every token!
      remove_padding = FALSE, # keep default, but doesn't matter since we didn't add padding when tokenizing
    verbose = quanteda_options("verbose")) |> # keep default
      #color is for least to most frequent
   textplot_wordcloud(max_words = 50,  color = c("red", "darkgreen", "blue")); title("FT: Pre")
```

```{r}
ft_post_toks |> dfm(tolower = TRUE, # yes, we want to lower case every token!
      remove_padding = FALSE, # keep default, but doesn't matter since we didn't add padding when tokenizing
    verbose = quanteda_options("verbose")) |> # keep default
      #color is for least to most frequent
   textplot_wordcloud(max_words = 50, color = c("red", "darkgreen", "blue")); title("FT: Post")
```

# CLOUDS FOR UK AND US TIMES
```{r}
library(quanteda.textplots)
us_pre_toks |> dfm(tolower = TRUE, 
      remove_padding = FALSE,
    verbose = quanteda_options("verbose")) |> 
   textplot_wordcloud(); title("US — Pre")

us_post_toks |> dfm(tolower = TRUE, 
      remove_padding = FALSE,
    verbose = quanteda_options("verbose")) |> 
   textplot_wordcloud(); title("US — Post")

uk_pre_toks |> dfm(tolower = TRUE,
      remove_padding = FALSE, 
    verbose = quanteda_options("verbose")) |>
   textplot_wordcloud(); title("UK — Pre") 

uk_post_toks |> dfm(tolower = TRUE, 
      remove_padding = FALSE, 
    verbose = quanteda_options("verbose")) |> 
   textplot_wordcloud(); title("UK — Post")

#Initial thoughts: UK pre has an emphasis on "people" and post had more on "company"
#us didn't change as much, ChatGPT became more in the headlines
```


# STRUCTURAL TOPIC MODELING: 
# - Like LDA but also takes into consideration that the articles differ between country and period 


```{r}
dfm_stm <- dfm_full |>
  dfm_trim(min_docfreq = 0.005, max_docfreq = 0.95, docfreq_type = "prop")

stm_in <- convert(dfm_stm, to = "stm")
# stm_in$documents, $vocab, $meta — meta carries your docvars
```


```{r}
set.seed(226)
fit <- stm::stm(
  documents  = stm_in$documents,
  vocab      = stm_in$vocab,
  K          = 20,                   
  prevalence = ~ period * country,   
  data       = stm_in$meta,
  init.type  = "LDA",         
  verbose    = TRUE
)

#Optimize the value of K, find the better value of K 
```


```{r}
lab <- labelTopics(fit, n = 7)

topic_table <- data.frame(
Topic = seq_len(nrow(lab$frex)),
FREX  = apply(lab$frex, 1, paste, collapse = ", "))
topic_table

#FREX brings up words that are frequent in a particular category but also exclusive to that category
```

